\section{Galerkin quadratures}
Until now, we have not described which angular quadrature should be used to
correctly treat high orders of anisotropy. We have only
stated that we need an appropriate quadrature but we did not explain what we
required. In this section, we introduce the Galerkin quadrature. Morel
first introduced them in \cite{galerkin_morel}; here, we
will introduce them following the presentation made in \cite{pautz_fp}. 

First, we start by recalling the definition of $\psi$ and $\phi_{l,m}$:
\begin{equation}
  \begin{split}
    \phi_{l,m}(\br) &= \int_{4\pi} d\bo' \psi(\br,\bo') Y_l^{m,*}\\
                    &=(\bs{D} \Psi)_{l,m}
  \end{split}
  \label{D2M}
\end{equation}
where $\bs{D}$ is the direction-to-moment operator. We also have:
\begin{equation}
  \begin{split}
    \Psi(\br,\bo) &= \sum_{l=0}^{\infty} \sum_{m=-l}^l Y_l^m(\bo)\phi_{l,m}(\br)\\
                  &= \bs{M}\Phi(\br)
  \end{split}
  \label{M2D}
\end{equation}
where $\bs{M}$ is the moment-to-direction operator. By combining \cref{D2M,M2D}, 
we obtain:
\begin{equation}
  (\bs{I}-\bs{MD})\Psi = 0
  \label{identity_alpha}
\end{equation}
since for analytic transport $\bs{M}=\bs{D}^{-1}$. $\bs{I}$ is the identity operator.

Now, we define \cite{pautz_fp}:
\begin{align}
\epsilon &= 1 - \la \bar{\mu}_0\ra \label{eps_pautz} \\
    \gamma &= \la\overline{(1-\mu_0)^2}\ra \label{gamma_pautz}\\
    A(\br) &= \frac{1-\bar{\mu}_0(\br)}{\epsilon}\label{A_pautz}\\
  \Sigma_a &= \hat{\Sigma}_a \label{hat_sig_a_0}\\
\Sigma_{s,l} &= \frac{\hat{\Sigma}_{s,l}(\br)}{\epsilon} \label{hat_sig_a_1}\\
\Sigma_{s,l}(\br) &=\Sigma_{s,0}(\br) \(1-\frac{l(l+1)}{2}A(\br)\epsilon +
  O(\gamma)\) \label{hat_sig_a_2}
\end{align}
where $\la X \ra$ is a typical value of $X$. Using \crefrange{eps_pautz}{hat_sig_a_2} 
in \cref{transport_sn}, we get:
\begin{equation}
\begin{split}
&\bo_k\cdot\bn\psi_k(\br) + \(\Sigma_a(\br)+\Sigma_{s,0}(\br)\) \psi_k(\br)
=\\
& \sum_{l=0}^{N-1} \sum_{m=-l}^l Y_l^m(\bo_k) \phi_{l,m}(\br)
\frac{\hat{\Sigma}_{s,0}(\br)}{\epsilon}\(1-\frac{l(l+1)}{2}A(\br)\epsilon +
O(\gamma)\) + Q(\br,\bo_k)
\end{split}
\label{discr_eps}
\end{equation}
where:
\begin{align}
\psi_k(\br) &= \psi(\br,\bo_k)\\
\phi_{l,m}(\br) &= \sum_{k=1}^K w_k Y_l^{m,*}(\bo_k) \psi_k(\br). \label{phi_p}
\end{align}
$w_k$ and $\bo_k$ are the quadrature weights and directions of a 
quadrature set of order $N$. For triangular quadrature sets, $K=N$ in 1D, 
$K=\frac{N(N+2)}{2}$ in 2D and $K=N(N+2)$ in 3D. \Cref{discr_eps} yields:
\begin{equation}
\begin{split}
&\bo_k\cdot\bn \psi_k(\br) +\hat{\Sigma}_a(\br)\psi_k(\br)+
\frac{\hat{\Sigma}_{s,0}(\br)}{\epsilon} \(\psi_k(\br)-\sum_{l=0}^{N-1}
\sum_{m=-l}^m  Y_l^m(\bo_k) \phi_{l,m}\)\\
&=-\frac{\(\Sigma_{tr}(\br)-\hat{\Sigma}_a(\br)\)}{2} \sum_{l=0}^{N-1}
\sum_{m=-l}^{l} l(l+1) Y_l^m(\bo_k) \phi_l^m(\br)\\
&+ Q(\br,\bo_k) + O\(\frac{\gamma}{\epsilon}\).
\end{split}
\label{discr_eps_manip}
\end{equation}
We insert the asymptotic ansatz:
\begin{align}
\psi &= \psi^{(0)} + \epsilon \psi^{(1)} + \epsilon^2\psi^{(2)}+\hdots\\
\phi_{l,m} &= \phi_{l,m}^{(0)} + \epsilon \phi_{l,m}^{(1)} + \epsilon^2
\phi_{l,m}^{(2)}+\hdots
\end{align}
into \cref{discr_eps}. The terms of $O(1)$ give:
\begin{equation}
\begin{split}
\phi_{l,m}^{(0)}(\br) &= \sum_{k=1}^K w_k Y_l^{m,*}(\bo_k) \psi_k^{(0)}(\br)\\ 
                      &= (\bs{D}_N\Psi^{(0)})_{l,m}.
\end{split}
\label{D2M_disc}
\end{equation}
Now we insert the ansatz into \cref{discr_eps_manip} and we look at the terms
of $O(\epsilon^{-1})$:
\begin{equation}
\begin{split}
\psi_d^{(0)}(\br) &=\sum_{l=0}^{N-1} \sum_{m=-l}^l Y_l^m(\bo_k) 
\phi_{l,m}^{(0)}(\br)\\
&= (\bs{M}_N\Phi^{(0)})_d
\end{split}
\label{M2D_disc}
\end{equation}
there is no $O(\gamma)$ term, since $\gamma\rightarrow 0$ when 
$\epsilon\rightarrow 0$, i.e., that there are no
$O(1)$ components in $\gamma$. \Cref{D2M_disc,M2D_disc} may be combined to give:
\begin{align}
  &(\bs{I}-\bs{M}_N\bs{D}_N)\Psi^{(0)} = 0 \label{identity_a}\\
  &(\bs{I}-\bs{D}_N\bs{M}_N)\Phi^{(0)} = 0 \label{identity_b}.
\end{align}
Therefore, $\Psi^{(0)}$ must be in the kernel of $\bs{I}-\bs{M}_N\bs{D}_N$ and
$\Phi^{(0)}$ must be in the kernel of $\bs{I}-\bs{D}_N\bs{M}_N$. Using 
\cref{phi_p}, \cref{identity_b} becomes successively:
\begin{equation}
  \(\bs{I}-\bs{D}_N \bs{M}_N\) \bs{D}_N \Psi^{(0)} = 0
\end{equation}
\begin{equation}
  \(\bs{D}_N-\bs{D}_N \bs{M}_N \bs{D}_N\) \Psi^{(0)} = 0
\end{equation}
\begin{equation}
  \bs{D}_N(\bs{I}-\bs{M}_N \bs{D}_N)\Psi^{(0)} = 0,
\end{equation}
which is always satisfied if \cref{identity_a} is satisfied.
Therefore, if \cref{identity_a} is satisfied, \cref{D2M_disc,M2D_disc} are
automatically satisfied.

A sufficient condition to satisfy \cref{identity_a} is that $\bs{M}_N \bs{D}_N
=\bs{I}$. This is of course true if $\bs{D}_N=\bs{M}_N^{-1}$ like in analytic 
transport. Obviously for $\bs{M}_N$ and $\bs{D}_N$ to be the inverse of each other, 
the matrices have to be square. Thus, the number of moments in the scattering 
expansion must be equal to number of discrete angles. In one-dimension,
$\bs{D}_N = \bs{M}_N^{-1}$ is satisfied if the 
quadrature set integrates exactly any polynomials of degree $2N-1$, like the 
Gauss-Legendre quadrature does. In multidimensional problems, the standard 
quadrature sets use more discrete angles than there are scattering moments. 
Therefore, $\bs{M}_N$ and $\bs{D}_N$ are rectangular matrices and they cannot 
be inverse of each other. In this case, \cref{identity_a} can only be satisfied if 
$\Psi^{(0)}$ is in the kernel of $(\bs{I}-\bs{M}_N\bs{D}_N)$. This can be achieved 
only if $\Psi^{(0)}$ satisfies nonphysical constraints \cite{pautz_fp}. If 
\cref{identity_a} is not satisfied, the asymptotic ansatz is not valid and 
there is no $O(1)$ solution to \cref{discr_eps,phi_p}.\\
If we assume that $\bs{M}_N \bs{D}_N=\bs{I}$, then the $O(\epsilon)$ terms in 
\cref{phi_p} give:
\begin{equation}
  \begin{split}
    \phi_{l,m}^{(1)}(\br) &= \sum_{k=1}^K w_k Y_{l}^{m,*} (\bo_k) \psi_k^{(1)}(\br)\\
                          &=(\bs{D}_N\Psi^{(1)})_{l,m}.
  \end{split}
\end{equation}
In \cref{discr_eps_manip}, the $O(1)$ terms give:
\begin{equation}
  \begin{split}
    &\bo_k\cdot \bn \psi_k^{(0)}(\br) + \hat{\Sigma}_a(\br)\psi_k^{(0)}(\br) +
    \hat{\Sigma}_{s,0}(\br) \(\psi_k^{(1)} (\br) - \sum_{n=0}^{N-1}
    \sum_{m=-l}^{l} Y_l^m (\bo_k)\phi_{l,m}^{(1)}(\br)\)\\
    &=-\frac{\Sigma_{tr}(\br)-\hat{\Sigma}_a(\br)}{2} \sum_{l=0}^{N-1}
    \sum_{m=-l}^l  l(l+1) Y_l^m(\bo_k)
    \phi_{l,m}^{(0)}(\br)+ Q(\br,\bo_k) + O\(\frac{\gamma}{\epsilon}\).
  \end{split}
  \label{O_1_terms}
\end{equation}
Now, we want the scattering term on the left side of \cref{O_1_terms} to
disappear in order to keep only $\psi_k^{(0)}$ and $\phi_{l,m}^{(0)}$. This 
is possible only if:
\begin{equation}
\psi_d^{(1),*}(\br) = \psi_d^{(1)}(\br),
\label{psi_1_star_old}
\end{equation}
where:
\begin{equation}
  \begin{split}
    \psi_d^{(1),*} &= \sum_{l=0}^{N-1}\sum_{m=-l}^l Y_l^m(\bo_k) 
    \phi_{l,m}^{(1)}(\br)\\
                   &= (\bs{M}_N\Phi^{(1)})_d\\
                   &=(\bs{M}_N \bs{D}_N \Psi^{(1)})_d,
  \end{split}
  \label{psi_1_star}
\end{equation}   
which is satisfied because of our previous assumption that $\bs{M}_N \bs{D}_N
= \bs{I}$. \Cref{O_1_terms} yields:
\begin{equation}
\begin{split}
  &\bo_k \cdot \bn \psi_k^{(0)}(\br)+\hat{\Sigma}_a(\br) \psi_k^{(0)}(\br) =
  \frac{\Sigma_{tr}(\br)-\hat{\Sigma}_a(\br)}{2}
  \(\(\frac{\partial}{\partial \mu}(1-\mu^2)\frac{\partial}{\partial
  \mu}+\right. \right.\\
  &\left.\left. \(\frac{1}{1-\mu^2}\)\frac{\partial^2}{\partial \varphi^2}\)
  \tilde{\psi}^{(0)} (\br,\bo)\)_{\bo=\bo_k}+ Q(\br,\bo_k) + 
  O\(\frac{\gamma}{\epsilon}\),
\end{split}
\label{last_pautz}
\end{equation}
where $k=1,\hdots,K$ and $\tilde{\psi}^{(0)}(\br,\bo)$ is an interpolant
through the points $P_k=\{\bo_k,\psi_k^{(0)}(\br,\bo)\}$. ``In one-dimensional 
geometry, $\tilde{\psi}^{(0)}(\br,\bo)$ is the $(N-1)-$order
polynomial interpolant through the points $P_k$. In multidimensional
geometries, $\tilde{\psi}^{(0)}(\br,\bo)$ is the spherical harmonic
interpolant through the points $P_k$'' \cite{pautz_fp}. A quadrature which 
satisfies the relation $\bs{D}_N = \bs{M}_N^{-1}$ is called a ``Galerkin'' quadrature 
because Morel derived it by using a Galerkin weighting method \cite{galerkin_morel}.

In \cite{galerkin_morel}, Morel made the following suggestions to find the 
correct limit using the $S_N$:
\begin{description}
\item [One-dimensional geometry:] The Gauss-Legendre quadrature set is the
  only quadrature set which is a Galerkin quadrature.
\item [Multidimensional geometry:] The standard quadrature sets have fewer 
  moments than discrete angles. Therefore to satisfy $\bs{D}=\bs{M}^{-1}$, spherical
  harmonics of higher order need to be added to increase the number of moments
  to equal the number of angular flux. Morel in \cite{galerkin_morel} and Reed in 
  \cite{reed} proposed an heuristic algorithm to choose the spherical 
  harmonics for multidimensional geometries. If we use the real spherical
  harmonics defined by:
  \begin{equation}
    Y_{l,m}^c = \left\{
      \begin{aligned}
        &Y_l^0 & \textrm{ if } m=0\\
        &\frac{1}{\sqrt{2}} \(Y_l^m+(-1)^{m} Y_l^{-m}\) & \textrm{ if } m> 0
      \end{aligned}
      \right.
  \end{equation}
  \begin{equation}
    Y_{l,m}^s = \frac{1}{i\sqrt{2}}(Y_l^{-m}-(-1)^{m}Y_l^m) \textrm{ if } m<0
  \end{equation}
  and a two-dimensional Cartesian geometry triangular quadrature set, the 
  selection rules to build $\bs{M}$ are the following:
  \begin{itemize}
    \item Take $Y_{l,m}^c$ if $l\in[0,N-1]$, $m\in[0,l]$, and $l+m$ is even.
    \item Take $Y_{l,m}^s$ if ($l \in [1,N-1]$, $m\in[1,l]$, and $l+m$ is
      even) or if ($l=N$, $m\in[1,l]$, and $l+m$ is even).
  \end{itemize}
  For three dimensional triangular quadrature sets, the spherical harmonics
  are:
  \begin{itemize}
    \item Take $Y_{l,m}^c$ if ($l \in [0,N-1]$ and $m\in[0,l]$) or if ($l =
      N$, $m\in[1,N-1]$, and $m$ is odd).
    \item Take $Y_{l,m}^s$ if ($l\in [1,N]$ and $m\in[1,l]$) or if ($l=N+1$,
      $m\in[2,N]$ and $m$ is even).
  \end{itemize}
\end{description}
General necessary conditions and explanations of the heuristic rules above
have been analyzed in \cite{galerkin_sanchez}.

Using a standard $S_n$ quadrature may lead to an unphysical solution for
charged particle transport. To further demonstrate the importance of Galerkin
quadrature, we define the scattering ratio matrix $\bs{C}$ by:
\begin{equation}
\bs{C} = \frac{1}{\Sigma_t}\bs{D}\bs{M}\bs{\Sigma},
\end{equation}
where $\bs{\Sigma}$ is the scattering matrix containing the moments of the
scattering cross sections on its diagonal.
$\bs{C}$ is a diagonal matrix whose entries are the scattering ratios
$c_l=\frac{\Sigma_{s,l}}{\Sigma_t} \leq 1$. The fact that $\bs{MD}=\bs{I}$
assures a one-to-one relation between the angular flux and the flux moments;
furthermore, the orthogonality of all spherical harmonic functions used in the
angular flux representation is preserved. However, if a standard quadrature is
employed, then $\bs{D} = \bs{M}^T \bs{W}$, where $\bs{W}$ is the diagonal
matrix of weights, and an inexact integration occurs for the set of spherical
harmonics than span $\bs{M}$. Thus, $\bs{C}$ will differ from
$\frac{\bs{\Sigma}}{\Sigma_t}$ and its eigenvalue could exceed unity. This is
equivalent to numerically adding multiplication into the medium and, depending
on the amount of leakage present in the geometrical configuration, the
steady-state transport equation may not reach a steady state solution
\cite{pautz_fp}.

A very important property of the Galerkin quadratures is that they treat 
delta function scattering exactly. This is very important for two reasons. 
Firstly, in charged-particle transport, some cross sections 
have the following form:
\begin{equation}
\Sigma^{k\rightarrow g}(\mu_0) = \Sigma^{k\rightarrow g} \delta(\mu_0-1),
\end{equation}
where $\Sigma^{k\rightarrow g}(\mu_0)$ is the differential cross section
associated with a transfer from group $k$ to group $g$. Thus, it is essential
that delta function scattering be treated exactly. Secondly and more
importantly since we work with energy-integrated equation, for 
electron scattering, the extended transport correction \cite{lathrop} can be 
used to reduce the within-group scattering cross sections by two orders of 
magnitude or more \cite{morel_79}. Without the extended transport correction, 
the scattering cross-section moments are very large and DSA does not 
accelerate the convergence of the solver. Because the Galerkin method treats 
delta function  scattering exactly, the extended transport correction does not
modify the solution of the problem \cite{morel_79}. This is very interesting 
since it allows significant reduction of the cross-section moments without
loss of accuracy. Showing this property is quite simple given what we have
presented thus far. We know that in the analytic case we have:
\begin{equation}
\int_{4\pi} \delta(\mu_0-1) \psi(\bo')\ d\bo' = \psi(\bo).
\end{equation}
When the scattering source is discretized, we have $\bs{\Sigma}=\bs{I}$.
Therefore, it is obvious that:
\begin{equation}
  \bs{D\Sigma M}\Psi = \Psi.
\end{equation}

To really understand the Galerkin quadrature, it is interesting to recall 
the development made in \cite{galerkin_morel}. First, we expand the
one-dimensional angular flux on an interpolatory trial space:
\begin{equation}
\Psi(\mu) = \sum_{d=1}^N \psi_d B_d(\mu).
\label{psi_b}
\end{equation}
Methods for generating the interpolatory basis function can be found in 
\cite{galerkin_morel}. Next, we expand the scattering source on the Legendre
polynomials:
\begin{equation}
\mc{S}(\mu) = \sum_{l=0}^{\infty} \frac{2l+1}{2} \xi_l P_l(\mu),
\label{S_mu_2}
\end{equation}
where:
\begin{equation}
  \xi_l = \int_{-1}^1 \mc{S}(\mu) P_l(\mu) d\mu.
\end{equation}
Now, we use the interpolatory trial space to approximate $S(\mu)$:
\begin{equation}
\tilde{\mc{S}}(\mu) = \sum_{d=1}^N \tilde{\mc{S}}_d B_d(\mu).
\label{tilde_s_mu}
\end{equation}
Since a Galerkin method is used, the residual associated with the trial space 
approximation must be orthogonal to the weighting
space. The residual associated with \cref{tilde_s_mu} is given by:
\begin{equation}
Res(\mu) = \tilde{\mc{S}}(\mu) - \mc{S}(\mu).
\label{R_mu}
\end{equation}
Since the Legendre polynomials form a basis for the weighting space, we
orthogonalize against the Legendre polynomials:
\begin{equation}
\int_{-1}^1 Res(\mu) P_l(\mu) d\mu = 0.0,\ \ l=0,\hdots,N-1.
\label{int_r_mu}
\end{equation}
\Cref{int_r_mu} is satisfied if:
\begin{align}
&\tilde{\xi}_l = \xi_l,\ \ l=0,\hdots,N-1 \label{xi_vs_xi}\\
&\tilde{\xi}_l = \int_{-1}^1 \tilde{\mc{S}}(\mu) P_l(\mu) d\mu.
\end{align}
The main idea of the Galerkin quadrature can be seen on \cref{xi_vs_xi}: ``the
discrete scattering source values are chosen such that the interpolatory
representation for that scattering source has the same Legendre moments
through degree $N-1$ as the exact scattering source calculated with the
interpolatory representation for the angular flux'' \cite{galerkin_morel}. 
Because all the elements of the weighting space can be expressed as a linear
combination of Legendre polynomials $P_l$ with $l<N$ and $\int_{-1}^{1}
P_l(x)P_m(x)\ dx=0$ for $l\neq m$, only the first $N-1$ cross-section moments
are needed.

Now, we compare the cross-section expansion for the Galerkin method and
the $S_n$ method. Both the Galerkin method and the standard $S_n$ method use
a Legendre expansion of the cross section. However, whereas the $S_n$ method
relies on the accuracy of the truncated Legendre expansion, the Galerkin 
scattering source is fully consistent with the exact cross section. It is not
important whether or not the truncated Legendre cross-section expansion 
accurately represents the scattering cross section. What is important is that
$\tilde{S}(\mu)$ is an accurate approximation of $S(\mu)$. This is a crucial 
property for charged particle transport. For example, the delta
function expansion of finite degree is never converged, and this is an issue if a 
$S_n$ method with a standard quadrature is used. However, as proved earlier, 
the scattering source computed by the Galerkin method is treated exactly.

If the scattering is isotropic, the $S_n$ method might appear superior to the
Galerkin, since the Galerkin requires the number of flux moments to be equal
to the number of angular flux. Fortunately, if the higher order cross-sections
moments are zero, only the first rows of $\bs{D}$ and the first columns of
$\bs{M}$ need to be kept.

A Galerkin quadrature set and its companion quadrature sets (standard
quadrature sets associated with the Galerkin quadrature set having weights
defined by the first row of the discrete-to-moment matrix) yield different
scattering only
if the expansion order of the scattering cross section is high enough. When the 
scattering is isotropic, the Galerkin quadrature set is the same as its
companion set. Even when the scattering is highly anisotropic, the results
given by the Galerkin set and its companion set can be very close. This is due
to the fact the high order moments of the scattering cross section are often,
but not always, very small. However, the results can also be quite different.
