%\chapter[ ]{\uppercase{ML options}}
\chapter{\uppercase{ML options}}

Some of the available coarsening schemes in ML \cite{ml_guide}:
\begin{description}
  \item[Uncoupled:] Attempt to construct aggregates of optimal size ($3^d$
    nodes in $d$ dimensions). Each process works independently and aggregates
    cannot span processes.
  \item[MIS:] Uses maximal independent set techniques to define aggregates.
    Aggregates can span processes. May provide better quality aggregates than
    {\bf Uncoupled}, but computationally more expensive because it requires
    matrix-matrix product.
\end{description}
Some of the smoothers:
\begin{description}
  \item[Jacobi]
  \item[Symmetric Gauss-Seidel]
\end{description}
Some of the coarse solvers:
\begin{description}
  \item[Jacobi]
  \item[Symmetric Gauss-Seidel]
  \item[Amesos-KLU:] Use {\bf KLU} through {\bf Amesos}. Coarse grid problem
    is shipped to processor 0, solved, and solution is broadcast.
  \item[Amesos-UMFPACK:] Use {\bf UMFPACK} through {\bf Amesos}. Coarse grid
    problem is shipped to processor 0, solved, and solution is broadcast.
  \item[Amesos-MUMPS:] Use double precision version of {\bf MUMPS} through
    {\bf Amesos}.
\end{description}
The MultiLevelPreconditioner class provide default values for five different
preconditioner types:
\begin{itemize}
  \item Classical smoothed aggregation for symmetric and positive definite or
    nearly symmetric and definite systems (used here)
  \item Classical smoothed aggregation-based two-level domain decomposition.
  \item Three-level algebraic domain decomposition.
  \item Eddy current formulation of Maxwell's equation.
  \item Energy-based minimizing smoothed aggregation suitable for highly
    convective nonsymmetric fluid flow problems.
\end{itemize}
The options used in this work are:
\begin{description}
  \item[option name:] SA
  \item[max levels:] 10
  \item[prec type:] $V-$cycle
  \item[aggregation type:] uncoupled-MIS
  \item[aggregation damping factor:] 4/3
  \item[eigen-analysis type:] cg
  \item[eigen-analysis iterations:] 10
  \item[smoother sweeps:] 2
  \item[smoother damping factor:] 1.0
  \item[smoother pre or post:] both
  \item[smoother type:] symmetric Gauss-Seidel
  \item[coarse type:] Amesos-KLU
  \item[coarse max size:] 128
\end{description}
