\chapter{\uppercase{Conclusions and future work}}\label{conclusion_chapter}
\section{Conclusions}
In this dissertation, we first recalled the development of the
Boltzmann-Fokker-Planck equation and the limitations of the Fokker-Planck operator.
The Boltzmann-Fokker-Planck equation was introduced for charged particle transport
because the scattering kernel is highly forward-peaked such that a standard Legendre
expansion of the scattering kernel is impractical and would require hundreds 
of terms. We also recalled a previous work from Pomraning showing that the 
Fokker-Planck operator is an asymptotic limit of the Boltzmann 
operator when the scattering is forward-peaked and that the energy transfer 
during a collision tends to zero. In the Boltzmann-Fokker-Planck equation, the 
Fokker-Planck operator is used to model the highly forward-peaked scattering
collisions whereas the Boltzmann operator is used for the wide angle
scattering collisions. The Fokker-Planck operator simplifies the calculation
of the transport equation but is valid only if the kernel is peaked enough.
For instance, realistic screened Rutherford cross sections are not peaked
enough. Then, we introduced the Fokker-Planck cross sections which mimic the
Fokker-Planck operator when using the Boltzmann operator. Since Fokker-Planck
cross sections are the most forward-peaked cross section (the extended
transport correction renders the delta scattering equivalent to no scattering
at all), we used them for our tests. Finally, we introduced Galerkin
quadratures. Galerkin quadratures are crucial to obtain correct results when
the scattering is highly anisotropic. Galerkin quadratures are equivalent to
standard quadratures when the scattering is weakly anisotropic but when the
scattering is highly anisotropic, using a standard quadrature can introduce
non physical solutions.

Next, we introduced the angular multigrid methods to speed up the calculation 
with highly anisotropic scattering. When the scattering is highly anisotropic, 
the Diffusion Synthetic Acceleration (DSA) is not effective anymore because it 
cannot speed up the convergence of high order flux moments. The initial work on 
this topic was carried out by Morel and Manteuffel. 
They developed a one-dimensional angular multigrid method  to
accelerate the convergence of Source Iteration (SI). This angular multigrid 
method uses an $S_{n/2}$ sweep to correct the $S_n$ sweep. The $S_{n/2}$ 
correction is itself corrected by a $S_{n/4}$ correction, etc. until the 
$S_4$ correction is corrected by $P1$ equations. They showed that when 
Fokker-Planck cross sections are used the spectral radius of the new 
method is bounded by 0.6 whereas the spectral radius of DSA can become 
arbitrary close to unity. Pautz, et al., generalized the angular multigrid 
method to multidimensional geometries. In this case, the successive correction
used an $S_2$ sweep before a DSA stage at the final level. Unfortunately, 
the generalized method was unstable. To stabilize it, the corrections need to 
be filtered with a diffusion operator. The effect of this diffusive filter 
is to remove the high frequencies of the correction. This method is better 
than straight SI+DSA but when the
anisotropy increases, the spectral radius can again become arbitrary close to one.
In this research, we recast the angular multigrid method for
multidimensional geometries without filtering as a preconditioner for a Krylov
subspace solver. The new method is always more effective and efficient than 
DSA and is more effective as the anisotropy increases but unlike the one 
dimensional method, the number of iterations does not saturate with 
increasing anisotropy.

We also extended the Modified Interior Penalty (MIP) DSA
developed for triangular cells to arbitrary polygonal meshes using the
PieceWise Linear Discontinuous (PWLD) finite elements. Arbitrary polygonal
meshes can potentially decrease the number of unknowns to mesh a domain and
they simplifies adaptive mesh refinement by suppressing hanging nodes. MIP
being SPD, is solved using CG. Therefore, we compared different
CG preconditioners: Algebraic MultiGrid (AMG) preconditioners and the more 
common Symmetric Gauss-Seidel (SGS) preconditioner. AMG were shown to be much 
more efficient than SGS. Among the different AMG methods tested the AGMG code 
was the fastest and about 20 times faster than CG used without preconditioning.

To end this conclusion, we recall the goals that we defined in the proposal:
\begin{itemize}
  \item reformulate the angular multigrid method as a preconditioner for
    Krylov solver for highly forward-peaked scattering.
  \item adapt MIP for quadrilateral and polygonal cells.
  \item test algebraic multigrid to solve efficiently MIP.
\end{itemize}
All these points have been treated in this dissertation.
